Our Galactocentric Coordinate System Is Probably Wrong, and Not by a Little

### 2018-12-11 paper/ms.tex
Update paper

\usepackage{graphicx}

@@

@@

Our analysis makes the assumption that a global midplane is well defined.
However, our results are also consistent with the notion that the global
midplane is not a well-defined concept. Furthermore, none of the galaxies chosen in this work have

@@

\section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
An analytic, axisymmetric potential with a known coordinate system is a far
too simple description of the Milky Way. We thus turn to cosmological zoom-in
simulations as approximations to a ``true'' Milky Way potential. We will
perform a few simulations of small clusters embedded in the potential of a
frozen $\z=0$ snapshot of one such zoom-in. We take our fiducial reference
frame from \citet{2018arXiv180610564S}.

Freezing the potential means that during the course of our simulation the
cluster will rapidly pass by much of the substructure in the potential. To
illustrate that this has little effect on our results, we resimulate each
cluster with the softening lengths of the embedded potential increased by a
factor of \Gus{something}.

Once we perform this simulation, we calculate the actions of each star in the
cluster assuming the best-fit smooth, axisymmetric potential, as we do
observationally for the Milky Way. We refer to this space as the observational
action space. We will find that there is significant time evolution in
observational action space. Attributing some of this evolution to our fiducial
coordinate system being incorrect, we minimize the time-dispersion in action
space

\subsection{Cluster Simulations} \label{ssec:cluster_sim}
We simulate a small star cluster of $256$ stars in the frozen $\z=0$
potential of m12i. We use the code \texttt{ph4} within the
\texttt{AMUSE} framework \citep{2011ascl.soft07007P, 2013CoPhC.184..456P,
2013A&A...557A..84P} for the cluster integration. We use a softening length
of $0.01\,\pc$ which means we do not correctly solve the internal dynamics of
the cluster, but this is not important for our purposes. To couple the
gravitational forces of the frozen snapshot to the cluster stars, we use the
\texttt{BRIDGE} \citep{2007PASJ...59.1095F}, also implemented in
\texttt{AMUSE}. Our simulation runs for $2\,\Gyr$, allowing us to solve for
$\sim 10$ orbits.

To compute the forces from the frozen potential, we use the simple tree code
\texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
at distances $R>50\,\kpc$ in the fiducial coordinate system.

We use the initial position and velocity of two star particles in
m12i. These two star particles are randomly chosen from all star
particles in the simulation with the following properties (in the fiducial
coordinate system):
\begin{enumerate}
    \item age in the range $(0.25, 0.75)\,\Gyr$,
    \item cylindrical radius in the range $(7.7, 8.7)\,\kpc$,
    \item height (absolute value) less than $0.5\,\kpc$,
    \item $J_r$ and $J_z$ in the range:
        \subitem $(10, 20)\,\actunit$ for the first star,
        \subitem $(40, 60)\,\actunit$ for the second star.
\end{enumerate}
The ids we choose are \texttt{23693026} for the first star and
\texttt{17012804} for the second star. We refrain from calling these ``low''
or ``high'' action stars since the actions are computed in a presumably
incorrect coordinate system.

To compute actions, both in the choosing of each of these star particles and
in what follows, we use the package
\texttt{agama}\footnote{\url{https://github.com/GalacticDynamics-Oxford/Agama},
commit: \texttt{0e48336}}. Using only the high-resolution particles, we
compute the symmetrized potential in two components. First, a multipole
expansion with $20$ radial grid points and a max order of $2$ for the
spherical-harmonic expansion is used for the dark matter. Second, a
cylindrical spline expansion with $20$ radial and $z$ grid points, a $0$ order
max index azimuthal-harmonic expansion, $0.2$ and $50\,\kpc$ for the minimum
and maximum radial grid node, $0.02$ and $10\,\kpc$ for the minimum and
maximum $z$ grid node is used for the baryonic matter (stars and gas). In both
cases, axisymmetry is explicitly enforced. Other \texttt{agama} parameters are
left as default.

\subsection{Frame Optimization} \label{ssec:frame_opt}
Figure~\Gus{something} shows the action computed at each time step in the
fiducial coordinate system. Clearly there is significant time evolution.

Even though we change the center, we do not recalculate the center velocity.
This is largely irrelevant, as the recalculated center velocity differs from
the fiducial center velocity by $<0.1\,\kms$. In recentering the galaxy, the
package \texttt{gizmo\_analysis} will subtract a slightly different value for
the Hubble flow from each particle's velocity to convert to a physical
coordinate system. The error this introduces is $\sim 0.02\,\kms$, which we
also ignore.

>>>>

\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{bm}

@@

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

@@

Our analysis makes the assumption that a global midplane is well defined.
However, our results are also consistent with the notion that the global
midplane is not a well-defined concept.

\subsection{Dynamical Midplane}
We now seek to define the midplane in a dynamically consistent manner, which
we refer to as the ``dynamical midplane''. By this, we mean we would like to
define the midplane in a manner consistent with some sort of dynamical
constraint. Our method, which we describe now, is just one of many possible
ways the dynamical midplane can be defined.

We define the dynamical midplane based on the principle that perfectly
circular, planar orbits will trace the midplane. Once the circular, planar
orbits have been identified, the procedure for finding the dynamical midplane
is relatively simple. From the measurement of each star's orbit, the midplane
of the galaxy can be computed at $z=0$ easily. Furthermore, we will find that
each orbit also gives a measure of the: dynamical center of the galaxy,
velocity of the galaxy, and acceleration of the galaxy.

We pause briefly to consider general difficulties in measuring stellar orbits
in a cosmological simulation. It may seem reasonable to place the galaxy in an
inertial reference frame and then to simply measure the positions of each star
particle over time. However, it is not generally the case that the center of
mass of the stellar disk experiences a vanishing acceleration. Indeed, in a
re-run of the simulations presented here\footnote{The re-run contained a
cosmic ray bug fix which is unimportant here, but the re-run also output the
acceleration of each particle, which is not the case for the simulations in
this work.}, we measure a total acceleration of $\sim 0.4\,\kms/\text{Myr}$
for stars with cylindrical radii within $0.5\,\kpc$ of $8.2\,\kpc$ and
vertical heights less than $0.4\,\kpc$. Assuming a constant acceleration, over
$22\,\Myr$ --- the time spacing of snapshots considered in this work --- this
results in a $\sim90\,\pc$ shift of the disk. Since the offset scales with
$t^2$, this presents a significant problem for measuring stellar orbits over
longer timescales.

Another option is to recenter and reorient the galaxy at each snapshot.
However, the centering and orientation procedures are not without
stochasticity themselves. This stochasticity can easily dominate the magnitude
of the orbit. For instance, a shift in the direction of the $z$-axis by a mere
$0.7\degree$ results in a shift in the midplane of $100\,\pc$ at a radius of
$8.2\,\kpc$. We attempted this procedure in various ways (i.e. exploring
different centering and orientation procedures), but were not able to cleanly
measure the orbits of disk stars.

Here we choose to simplify the problem by considering only perfectly circular,
planar orbits. Since these orbits have no phase-dependent structure, we can
work on shorter timescales, i.e. for orbits with significant radial or
vertical action a large fraction of their orbit would need to be sampled in
order to measure the orbit accurately. Because we know the analytic form of
these orbits, we can forward model the trajectory of each star particle in the
cosmological simulation. Working in comoving coordinates, we can write down
the trajectory of a star particle as

\beq
\bm{r}(t) = R(\theta, \phi, \psi) \left( (A\cos{\omega t}, B\sin{\omega t}, C t) + \bm{x_0} \right)
\eeq

@@

% \section{Optimally Axisymmetric Frame} \label{sec:oa_frame}
% An analytic, axisymmetric potential with a known coordinate system is a far
% too simple description of the Milky Way. We thus turn to cosmological zoom-in
% simulations as approximations to a ``true'' Milky Way potential. We will
% perform a few simulations of small clusters embedded in the potential of a
% frozen $\z=0$ snapshot of one such zoom-in. We take our fiducial reference
% frame from \citet{2018arXiv180610564S}.

% Freezing the potential means that during the course of our simulation the
% cluster will rapidly pass by much of the substructure in the potential. To
% illustrate that this has little effect on our results, we resimulate each
% cluster with the softening lengths of the embedded potential increased by a
% factor of \Gus{something}.

% Once we perform this simulation, we calculate the actions of each star in the
% cluster assuming the best-fit smooth, axisymmetric potential, as we do
% observationally for the Milky Way. We refer to this space as the observational
% action space. We will find that there is significant time evolution in
% observational action space. Attributing some of this evolution to our fiducial
% coordinate system being incorrect, we minimize the time-dispersion in action
% space

% \subsection{Cluster Simulations} \label{ssec:cluster_sim}
% We simulate a small star cluster of $256$ stars in the frozen $\z=0$
% potential of m12i. We use the code \texttt{ph4} within the
% \texttt{AMUSE} framework \citep{2011ascl.soft07007P, 2013CoPhC.184..456P,
% 2013A&A...557A..84P} for the cluster integration. We use a softening length
% of $0.01\,\pc$ which means we do not correctly solve the internal dynamics of
% the cluster, but this is not important for our purposes. To couple the
% gravitational forces of the frozen snapshot to the cluster stars, we use the
% \texttt{BRIDGE} \citep{2007PASJ...59.1095F}, also implemented in
% \texttt{AMUSE}. Our simulation runs for $2\,\Gyr$, allowing us to solve for
% $\sim 10$ orbits.

% To compute the forces from the frozen potential, we use the simple tree code
% \texttt{pykdgrav}\footnote{\url{https://github.com/omgspace/pykdgrav},
% commit: \texttt{fbd1ddc}}. We use an opening angle of $0.5$ and exclude stars
% at distances $R>50\,\kpc$ in the fiducial coordinate system.

% We use the initial position and velocity of two star particles in
% m12i. These two star particles are randomly chosen from all star
% particles in the simulation with the following properties (in the fiducial
% coordinate system):
% \begin{enumerate}
%     \item age in the range $(0.25, 0.75)\,\Gyr$,
%     \item cylindrical radius in the range $(7.7, 8.7)\,\kpc$,
%     \item height (absolute value) less than $0.5\,\kpc$,
%     \item $J_r$ and $J_z$ in the range:
%         \subitem $(10, 20)\,\actunit$ for the first star,
%         \subitem $(40, 60)\,\actunit$ for the second star.
% \end{enumerate}
% The ids we choose are \texttt{23693026} for the first star and
% \texttt{17012804} for the second star. We refrain from calling these ``low''
% or ``high'' action stars since the actions are computed in a presumably
% incorrect coordinate system.

% To compute actions, both in the choosing of each of these star particles and
% in what follows, we use the package
% \texttt{agama}\footnote{\url{https://github.com/GalacticDynamics-Oxford/Agama},
% commit: \texttt{0e48336}}. Using only the high-resolution particles, we
% compute the symmetrized potential in two components. First, a multipole
% expansion with $20$ radial grid points and a max order of $2$ for the
% spherical-harmonic expansion is used for the dark matter. Second, a
% cylindrical spline expansion with $20$ radial and $z$ grid points, a $0$ order
% max index azimuthal-harmonic expansion, $0.2$ and $50\,\kpc$ for the minimum
% and maximum radial grid node, $0.02$ and $10\,\kpc$ for the minimum and
% maximum $z$ grid node is used for the baryonic matter (stars and gas). In both
% cases, axisymmetry is explicitly enforced. Other \texttt{agama} parameters are
% left as default.

% \subsection{Frame Optimization} \label{ssec:frame_opt}
% Figure~\Gus{something} shows the action computed at each time step in the
% fiducial coordinate system. Clearly there is significant time evolution.

% Even though we change the center, we do not recalculate the center velocity.
% This is largely irrelevant, as the recalculated center velocity differs from
% the fiducial center velocity by $<0.1\,\kms$. In recentering the galaxy, the
% package \texttt{gizmo\_analysis} will subtract a slightly different value for
% the Hubble flow from each particle's velocity to convert to a physical
% coordinate system. The error this introduces is $\sim 0.02\,\kms$, which we
% also ignore.
